---
title: "Political party colours"
author: "Zoe Meers"
date: "14/10/2019"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = FALSE,
  echo = FALSE
)
library(tidyverse)
library(ussc)
ussc_fonts()
library(gt)
```

Sometimes you want to colour points or lines by political parties i.e. Republicans and Democrats or ALP/Coalition/Greens. To make this easy to do, I have stored party information in `ussc_colours`. To access the colours as a vector, run `ussc_pal$auspol` or `ussc_pal$uspol`.  If you want to use the colours in a `ggplot2` object, simply create your graph and append `scale_colour_ussc("auspol")` or `scale_fill_ussc("auspol")` to the object. 

```{r, echo = TRUE}
label <- read.csv("candidates_midterms.csv")%>%
  filter(Cycle == "2018")

read.csv("candidates_midterms.csv") %>%
  ggplot() +
  geom_line(aes(Cycle, Female.Candidates, colour = party), size = 1.5) +
  geom_point(aes(Cycle, Female.Candidates, colour = party), size = 2.5) +
  theme_ussc() +
  directlabels::geom_dl(
    data = label, aes(Cycle, Female.Candidates,
      label = paste0(party, " (", Female.Candidates, ")"),
      colour = party
    ),
    method = list(directlabels::dl.trans(x = x - 3.5, y = y + .1),
      "last.points",
      cex = 1
    )
  ) +
  scale_colour_ussc("uspol") +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  ) +
  scale_x_continuous(breaks = seq(1994, 2018, 2)) +
  labs(
    title = "Women candidates in the United States",
    y = "", colour = "", x = "",
    caption = "Source: Centre for Responsive Politics"
  )