---
title: "Roll Call Data & Ideal Point Estimates"
author: "Zoe Meers"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# D3 in R

RStudio recently launched `R2D3`, giving `R` users the ability to render `D3.js` code in RStudio. For those in the data visualization field, this is a *really* exciting development. Now we can render `D3` code in `RMarkdown`, `R` script, or `Shiny`. We can also pass `R` objects (i.e. dataframes, datatables, etc.) to `D3` visualizations.

Secondly, we can wrap `r2d3` functions and include them in packages! As an example, I've taken Simon's ideal point estimates `D3` code, rewrote it so that it would work in `r2d3`, wrapped the result up into a function called `d3_rollcall_idealpoints()` and uploaded it to the `ussc` `R` package. 
I'll walk you through the steps to run the function. Feel free to check out the [`r2d3` website](https://rstudio.github.io/r2d3) if you want to learn more about `D3` in `R`.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings=FALSE, message=FALSE)
library(r2d3)
library(here)
library(ussc)
#set wd or open the file directly. 
```


```{r}
#load data
house_estimates <- read.csv(here("h_estimates.csv"))
head(house_estimates)
```
* Note: this code will really only work with Simon's ideal points data as the defined variables in the D3 code are the same as this csv file. I would like to make a generic ideal points function in the future but this may be a while off.


## 115th House of Representatives Ideal Point Estimates
```{r, fig.height=20}
#Run d3_rollcall_idealpoints for House of Rep data
#You *must* define the data frame -- we did that above, reading the csv file to house_estimates and defined the data below in the function itself. This is the beauty of r2d3 -- any data frame or R object will work if the variables match.
# Note: you can also set the height and width of the visualization inside the function or you can also that in the R chunk.
d3_rollcall_idealpoints(data=house_estimates)
```

```{r}
#load data
senate_estimates <- read.csv(here("s_estimates.csv"))
```

## 115th Senate Ideal Point Estimates
```{r, fig.height=5}
#run d3_rollcall_idealpoints for Senate data
d3_rollcall_idealpoints(data=senate_estimates)
```

There you have it! Pretty simple, huh? Interactive `D3.js` visualizations without the hassle. 