---
title: "`theme_ussc_dark()`"
author: "Zoe Meers"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = FALSE
)
library(ggplot2)
library(ussc)
```

## Setting the background colour of a `ussc` `ggplot2` graph

All `ussc` themes have a transparent background which means that you can set the background colour of a graph by using developer arguments. To do this, append a CSS style sheet in the R Markdown YAML (which affects the entire HTML document) or write ```{r dev.args = list(bg = 'colour')}``` for a specific graph, replacing colour with a HEX code or rgb code. Note that if you do not set a background colour, `theme_ussc()` and other variations in the `ussc` package will default to a white background.

### Example 1

#### ```{r fig.align='center', dev.args=list(bg='#2a2a2b')}```

```{r univers ussc theme and blue colour scale, fig.align='center', dev.args=list(bg='#2a2a2b')}
ggplot(iris, aes(Sepal.Width, Sepal.Length, colour = Species)) + 
  geom_point(size = 4, alpha=0.4) +  
  theme_ussc_dark() + 
  labs(title="Univers Header", 
       x="Univers Font: Sepal Width", 
       y="Univers Font: Sepal Length", 
       caption = "The data is from the Iris R package and is an example.") + 
  scale_colour_ussc("blue", reverse=TRUE)
```

### Example 2

#### ```{r GAM, fig.align='center', dev.args=list(bg='lightblue')}```

```{r GAM, fig.align='center', dev.args=list(bg='lightblue')}
ggplot(data=iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species)) + 
  geom_point(aes(shape=Species), size=1.5) + 
  xlab("Sepal Length") + 
  ylab("Sepal Width") + 
  ggtitle("Scatterplot with smoothers") + 
  scale_colour_ussc("main") + 
  theme_ussc() + 
  geom_smooth(method="gam", formula= y~s(x, bs="cs"))
```
