---
title: "USSC Confluence functions"
author: "Zoe Meers"
date: "19/03/2019"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = FALSE
)
library(ggplot2)
library(tidyverse)
library(fs)
library(ussc)
```

# Confluence REST API

The US Studies Centre has an online collaboration software called Confluence. It was developed by Atlassian, the technology company. Confluence has a REST API which means we can directly load data from Confluence into `R`. 

`ussc` provides two functions that will be useful for researchers: `ussc_confluence_table()` and `ussc_confluence_excel()`.

To use these functions, you need to connect to the REST API. Grab an API key from https://confluence.atlassian.com/cloud/api-tokens-938839638.html. You can add the API key into your .renviron file by running `usethis::edit_r_environ()`. The confluence keys look like this in my .renviron file: 

CONFLUENCE_URL = https://usscsydney.atlassian.net/wiki

CONFLUENCE_USERNAME = firstname.lastname@sydney.edu.au (add your own email)

CONFLUENCE_PASSWORD = KFkuYkNdkqSKEWfdrjODxxxx (change to your API key -- note that this API key does not work as I changed the last four digits)


Save the .renviron file and restart `R` so the changes take effect. 


## Scrape HTML tables from Confluence

`ussc_confluence_table()` returns HTML tables that have been embedded in Confluence as tidy tibbles. 
There are three parameters - the ID of the Confluence page, your username and your password (API key). The username and password default to CONFLUENCE_USERNAME and CONFLUENCE_PASSWORD in your .renviron file. You can add the values manually if you wish.

It is a fairly simple function to run. Say I want to grab the Key Performance Indicators from the Comms team. The page ID is 950239240.

```{r}
comms_kpi <-  ussc_confluence_table("950239240")
```

This grabs all of the tables on the page. You can extract each table by running: 

```{r}
online_kpi_table1 <- comms_kpi[1][[1]]

head(online_kpi_table1)
```


## Download Excel files from Confluence

The second function, `ussc_confluence_excel()`, downloads all Excel files on a given page. Note: this function might take a while to run if you need to download many files. I have only tested this function on Firefox. It also assumes that you have set your settings so that the files download automatically.

```{r}
one_file <- ussc_confluence_excel(id = "950239621")

```

The Excel file has 4 sheets. We can extract the sheets below.

```{r}
publications <- one_file[["Publications"]]
media <- one_file[["Media"]]
events <- one_file[["Events"]]
meetings_etc <- one_file[["Meetings and other"]]


head(publications)
```


