---
title: "USSC Confluence functions"
author: "Zoe Meers"
date: "19/03/2019"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = FALSE
)
library(ggplot2)
library(tidyverse)
library(fs)
library(ussc)
```

# Confluence REST API

`ussc` provides five functions that will be useful if/when you need to download data from Confluence: 

* `ussc_confluence_table()` to grab any table and return in the same format as on Confluence, 
* `ussc_confluence_excel()` for downloading Excel files and loading into `R`,
* `ussc_confluence_kpi_table()` for grabbing the KPIs and cleaning the data so it is long and tidy,
* `ussc_confluence_version_history()` to grab the version history of any table on Confluence (particularly useful for updates to the Publications calendar)
* `ussc_confluence_word_tables()` to download tables within Word documents, load them into `R` and clean them.

To use these functions, you need to connect to the Confluence REST API. Grab an API key from https://confluence.atlassian.com/cloud/api-tokens-938839638.html. You can add the API key into your .renviron file by running `usethis::edit_r_environ()`. The confluence keys look like this in my .renviron file: 

CONFLUENCE_URL = https://usscsydney.atlassian.net/wiki

CONFLUENCE_USERNAME = firstname.lastname@sydney.edu.au (add your own email)

CONFLUENCE_PASSWORD = KFkuYkNdkqSKEWfdrjODxxxx (please change your API key -- note that this API key does not work as I changed the last four digits)


Save the .renviron file and restart `R` so the changes take effect. 


## Scrape HTML tables from Confluence

`ussc_confluence_table()` returns HTML tables that have been embedded in Confluence as tidy tibbles. 
There are three parameters - the ID of the Confluence page, your username and your password (API key). The username and password default to CONFLUENCE_USERNAME and CONFLUENCE_PASSWORD in your .renviron file. You can add the values manually if you wish.

It is a fairly simple function to run. Say I want to grab the Key Performance Indicators from the Comms team. The page ID is 950239240.

```{r}
comms_kpi <-  ussc_confluence_table("950239240")
```

This grabs all of the tables on the page. You can extract each table by running: 

```{r}
online_kpi_table1 <- comms_kpi[1][[1]]

head(online_kpi_table1)
```
## Publications KPI table

If you want to grab the publication metrics data, run `ussc_confluence_kpi_table()`. This will download the data and clean it. Let's take a look at the 2020 Q1 metrics.

```{r}
pub_kpi <- ussc_confluence_kpi_table("1173357736")

head(pub_kpi)
```

## Version history

To see the version history of a table, run the function `ussc_confluence_version_history()`. You will be able to see when edits to the page were made and by whom.

```{r}
pubs_calendar_version_history <- ussc_confluence_version_history("1020526593")

head(pubs_calendar_version_history)
```



## Download Excel files from Confluence

Another function, `ussc_confluence_excel()`, downloads all Excel files on a given page. Note: this function might take a while to run if you need to download many files. I have only tested this function on Firefox. It also assumes that you have set your settings so that the files download automatically.

```{r}
one_file <- ussc_confluence_excel(id = "950239621")
```

The Excel file has 4 sheets. We can extract the sheets below.

```{r}
publications <- one_file[["Publications"]]
media <- one_file[["Media"]]
events <- one_file[["Events"]]
meetings_etc <- one_file[["Meetings and other"]]


head(publications)
```


