---
title: "Visual explanation of USSC ggplot2 themes and colour scales"
author: "Zoe Meers"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warnings=FALSE,
  message = FALSE,
  comment = FALSE
)
```

# Introduction

This vignette visually explains how researchers at the USSC can use the Center's fonts and colours in their research output. 

# Setup
In addition to loading the packages you would otherwise use in your analysis, load the USSC package. In this example, I have loaded tidyverse along with ussc.

```{r}
library('ussc')
library('tidyverse')
```

# Example 1

The first example shows the standard ussc theme. The font used in the title and caption is Neo Sans Pro. In other cases, theme_ussc() defaults to Univers (see axes and legend). In this scatterplot, I have defined the colour using scale_colour_ussc(). I call the main colour palette which is an interpolation scale ranging from the dark blue colour scheme to red. In this example, the colour scale has been reversed from red to dark blue.
```{r ussc theme and main colour scale, fig.align='center', fig.width=6, fig.height=5}
ggplot(iris, aes(Sepal.Width, Sepal.Length, colour = Species)) + 
  geom_point(size = 4, alpha=0.4) +  
  theme_ussc() + 
  labs(title="Neo Sans Pro Header", x="Univers Font: Sepal Width", y="Univers Font: Sepal Length", caption = "The data is from the Iris R package and is an example.") + 
  scale_colour_ussc("main", reverse=TRUE) 
```

# Example 2

Theme_ussc_univers is a wrapper for theme_bw() that uses the Univers font. We have set the scale_colour_ussc to the blue palette which ranges from the official USSC dark blue colour to light blue.
```{r univers ussc theme and blue colour scale, fig.align='center', fig.width=6, fig.height=5}
ggplot(iris, aes(Sepal.Width, Sepal.Length, colour = Species)) + 
  geom_point(size = 4, alpha=0.4) +  
  theme_ussc_univers() + 
  labs(title="Univers Header", x="Univers Font: Sepal Width", y="Univers Font: Sepal Length", caption = "The data is from the Iris R package and is an example.") + 
  scale_colour_ussc("blue", reverse=TRUE)
```

# Example 3

This chart uses the light Univers font and the main colour palette. The scatter points are opaque (i.e. I have not used alpha to increase transperancy).
```{r univers light theme and main colour scale w/o alpha, fig.align='center', fig.width=6, fig.height=5}
ggplot(iris, aes(Sepal.Width, Sepal.Length, colour = Species)) + 
  geom_point(size = 4) +  
  theme_ussc_univers_light() + 
  labs(title="Univers Header", x="Univers Font: Sepal Width", y="Univers Font: Sepal Length", caption = "The data is from the Iris R package and is an example.") + 
  scale_colour_ussc("main") 
```

# Example 4

This chart uses the light Neo Sans Pro font and the black and white colour palette. Alpha has been set to 70%.

```{r neo sans pro light theme and grey colour scale, fig.align='center', fig.width=6, fig.height=5}
ggplot(iris, aes(Sepal.Width, Sepal.Length, colour = Species)) + 
  geom_point(size = 4, alpha=0.7) +
  theme_ussc_neosanspro_light() + 
  labs(title="Neo Sans Pro Header", x="NSP Font: Sepal Width", y="NSP Font: Sepal Length", caption = "The data is from the Iris R package and is an example.") + 
  scale_colour_ussc("grey")
```

# Example 5

Here is an example where the fill is discrete and the graph has been facetted into a grid. 
```{r facet ussc theme and main colour scale, fig.align='center', fig.width=6, fig.height=5}
ggplot(mtcars, aes(mpg, hp, colour=mpg)) + 
  geom_point() + 
  facet_grid(cyl~gear) + 
  theme_ussc() + 
  scale_fill_ussc("main", discrete=TRUE) + labs(title='A facet_grid example')
```


