---
title: "`theme_ussc_dark()`"
author: "Zoe Meers"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = FALSE
)
```


```{r}
library('ussc')
library('tidyverse')
```

# Introduction

This vignette is a quick introduction to using the dark USSC theme. 

The dark USSC theme has a transparent background, meaning that the user needs to set the background colour for the code chunk or for the HTML themselves. (Actually, all of our themes have a transparent background but you wouldn't know this if you're plotting a regular graph with a white background.)

The way to do this is to set the `dev.args` call in the code chunk by writing `{r dev.args = list(bg = 'color')}`:

```{r dev.args = list(bg = '#2a2a2b'), fig.width = 7, fig.height = 7}
ggplot(iris, aes(Sepal.Width, Sepal.Length, colour = Species)) + 
  geom_point(size = 4, alpha=0.4) +  
  theme_ussc_dark() + 
  labs(title="Univers Header", x="Univers Font: Sepal Width", y="Univers Font: Sepal Length", caption = "The data is from the Iris R package and is an example.") + 
  scale_colour_ussc("main", reverse=TRUE) 
```

```{r GAM, fig.align='center', dev.args = list(bg = 'lightblue'), fig.width = 7, fig.height=7}
ggplot(data=iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species)) + 
  geom_point(aes(shape=Species), size=1.5) + 
  xlab("Sepal Length") + 
  ylab("Sepal Width") + 
  ggtitle("Scatterplot with smoothers") + 
  scale_colour_ussc("main") + 
  theme_ussc() + 
  geom_smooth(method="gam", formula= y~s(x, bs="cs"))

ggplot(data=iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species)) + 
  geom_point(aes(shape=Species), size=1.5) + 
  xlab("Sepal Length") + 
  ylab("Sepal Width") + 
  ggtitle("Scatterplot with smoothers") + 
  scale_colour_ussc("main") + 
  theme_ussc_dark() + 
  geom_smooth(method="gam", formula= y~s(x, bs="cs"))
```
